Barneby Lives usage example

### 2023-05-17 scripts/project_template.Rmd
converting date functions to tidy

```{r}
dummy_pts <- data.frame(
  collectionNo = paste0(rep('RCB', 1000), seq(1:1000)),
  day = c(sample(1:31, size = 1000, replace = T)),
  month = c(sample(5:10, size = 1000, replace = T)),
  year = 2023,
  x = runif(1000, min = -120, max = -100), 
  y = runif(1000, min = 35, max = 48)
) %>% sf::st_as_sf(coords = c(x = 'x', y = 'y'), crs = 4326)

@@

out1 <- political_grabber(dummy_pts, 'collectionNo')
out2 <- place_grabber(dummy_pts)
out3 <- physical_grabber(dummy_pts)
```

>>>>

```{r dummy data}
dummy_pts <- data.frame(
  collectionNo = paste0(rep('RCB', 1000), seq(1:1000)),
  day = c(sample(1:31, size = 1000, replace = T)),
  month = c(sample(5:10, size = 1000, replace = T)),
  year = 2023,
  x = runif(1000, min = -120, max = -100), 
  y = runif(1000, min = 35, max = 48)
) %>% sf::st_as_sf(coords = c(x = 'x', y = 'y'), crs = 4326)

 first convert from american date format (m%d%year%) to ISO 8601 for both collection and determination date

 second from ISO 8601 prepare notation dates for day of collection and date of determination in text format

```{r Date parsing}

#' create herbarium format dates for specimens
#' 
#' this function will return up to two additional columns with x. 'coll_date_text' a text format for the date of collection and 'det_date_text" a shorter text format date for identification
#' @param x an (sf/tibble/) data frame with both the collection date
#' and determination date
#' @param coll_date date of collection, expected to be of the format 'MM/DD/YYYY', minor checks for compliance to the format will be carried out before returning an error.
#' @param det_date date of determination, same format and processes as above. 
#' @returns coll_date ISO 8601, det_date ISO 8601, month, day, year, coll_date_text, det_date_text columns for printing onto the label
date_parser <- function(x, coll_date, det_date){

  coll_date_q <- enquo(coll_date)
  det_date_q <- enquo(det_date)

  mdy2dmy <- function(x){format(as.Date(x, format = "%m/%d/%Y"), "%d/%m/%Y")}

  x_dmy <- x |> 
    mutate(across(.cols = c(!!coll_date_q, !!det_date_q), mdy2dmy,
           .names = "{col}_dmy"), .before = ncol(.)) |>
    mutate(across(.cols = c(!!coll_date_q, !!det_date_q), ~ lubridate::month(x),
           .names = "{col}_mo"))

  # add in collection dates here
#    if(missing(det_date)) {
#      with_date <- dplyr::mutate(x_dmy, coll_date_text = date2text(x_dmy, 'coll_date_dmy'))
##    } else {
#     with_date <-  dplyr::mutate(x_dmy, 
#           coll_date_text = date2text(x_dmy, 'coll_date_dmy'),
#           det_date_text  = date2text(x_dmy, 'coll_date_dmy'))
#    }

  return(x_dmy)
}

```{r}

fourtet <- data.frame(col_date = c('12/23/2003','07/09/2023', '1/13/2016'),
                   #   det_date = c('12/25/2003','07/09/2023', '1/13/2016'),
                      tet = c('abc', 'def', 'ggg'))

date_parser(fourtet, coll_date = 'col_date')

@@

out1 <- political_grabber(dummy_pts, 'collectionNo')
out2 <- place_grabber(dummy_pts)
out3 <- physical_grabber(dummy_pts)

browseVignettes("ggtree")
```
